\section{Introduction}
\label{sec:ch1}

Classical Ramsey Theory asks for the minimum number of vertices required to guarantee the existence of a clique or an independent set in a graph. We refer to these maxima as \textit{Ramsey numbers}.

\begin{definition}
  \label{def:ramsey-number}
  The \textit{Ramsey number} $R(s, t)$ is the minimum number $N$ such that any graph on $N$ vertices contains a clique of size $s$ or an independent set of size $t$.
\end{definition}

The existence of these numbers was proven by Ramsey~\cite{ramsey1930} in 1930, but their quantitative behaviours remain largely unknown despite decades of intensive research. In particular, it is known that $R(3, 3) = 6$ and $R(4, 4) = 18$, but the exact values of $R(5, 5)$ and beyond are yet to be determined. The cornerstone of Ramsey Theory due to Erd\H{o}s and Szekeres in 1935 gives the following upper bound on $R(s, t)$:
\begin{theorem}[Erd\H{o}s--Szekeres~\cite{erdos1935some}]
  \label{thm:erdos-szekeres}
  For $s, t \geq 2$,
  \[
    R(s, t) \leq \binom{s + t - 2}{s - 1}.
  \]
\end{theorem}
Since $\binom{2t-2}{t-1} < 2^{2t}$, we get $R(t,t) < 4^t$. This bound resisted improvement until the recent breakthrough of Campos, Griffiths, Morris, and Sahasrabudhe~\cite{campos2025exponentialimprovementdiagonalramsey}, giving the first exponential improvement since Erd\H{o}s--Szekeres. The lower bound of Ramsey numbers has been even harder to determine. A rather basic result which already gives the best known exponents is the famous random graph construction due to Erd\H{o}s in 1947, marking one of the first applications of the probabilistic method in combinatorics. Before stating the result, we define the \textit{Erd\H{o}s--R\'{e}nyi random graph}.

\begin{definition}
  \label{def:random-graph}

  For $n \geq 1$ and $p \in [0, 1]$, the \textit{random graph} $G_{n, p}$ is the graph on $n$ vertices in which each of the $\binom{n}{2}$ edges is present independently with probability $p$.
\end{definition}

\textit{Remark: For constant $p$, $G_{n, p}$ is dense in the sense that it has $\Theta(n^2)$ many edges almost surely.}

\begin{theorem}[Erd\H{o}s \cite{erdos1972ramsey}]
  \label{thm:diagonal-ramsey}
  For $t \geq 3$,
  \[
    \sqrt{2}^t < R(t, t) < 4^t.
  \]
\end{theorem}

\begin{proof}
  The upper bound follows from Theorem~\ref{thm:erdos-szekeres}. For lower bound, set $N = \lfloor \sqrt{2}^t \rfloor$ and consider the random graph $G_{N, 1/2}$. The expected number of cliques of size $t$ is
  \[
    \binom{N}{t} \cdot 2^{-\binom{t}{2}} < \frac{N^t}{t!}2^{-\binom{t}{2}} \leq \frac{\sqrt{2}^{t^2}}{t!}2^{-\binom{t}{2}} = \frac{\sqrt{2}^t}{t!} = \frac{\sqrt{2}^{t - 3}}{t(t - 1)\cdots 4} \cdot \frac{\sqrt{2}^3}{3!} \leq \frac{\sqrt{2}^3}{3!} < \frac{1}{2}.
  \]
  Similarly, the expected number of independent sets of size $t$ is less than $1/2$. It now follows that there exists a graph on $N$ vertices containing neither a clique nor an independent set of size $t$.
\end{proof}

% The above strategy may be optimized to give a lower bound of $\frac{{t}}{e\sqrt{2}}\sqrt{2}^t$, which has been astonishingly resilient to improvement. The only noteworthy improvement since 1947 is achieved in 1977 by Spencer~\cite{Spencer1975RamseysT} using the Lov\'{a}sz local lemma to raise the bound to $(1 - o(1))\frac{t\sqrt{2}}{e}\sqrt{2}^t$, an improvement by a factor of 2. For further background on Ramsey theory, see Conlon, Fox, and Sudakov~\cite{conlon2015recentdevelopmentsgraphramsey}.

We define a shorthand for cliques and independent sets.

\begin{definition}
  \label{def:hom}

  A set $S \subseteq V(G)$ is \textit{homogeneous} in a graph $G$ if it induces a clique or an independent set in $G$. We denote the size of the largest homogeneous set in $G$ by $\hom(G)$.
\end{definition}

Therefore, Theorem~\ref{thm:diagonal-ramsey} may be restated as
\[
  \frac{1}{2}\log_2 n < \hom(G) < 2\log_2 n,
\]
for any graph $G$ on $n$ vertices. Apart from determining the bounds of Ramsey numbers, the search for explicit graph constructions without large homogeneous sets has also been a major open problem in Ramsey Theory. We refer to such graphs as \textit{Ramsey graphs}.
\begin{definition}
  \label{def:ramsey-graph}

  Let $C > 0$ be a constant. A \textit{$C$-Ramsey graph} is a graph with no homogeneous set of size greater than $C\log n$.
\end{definition}

\textit{Remark: } For constant $p \in (0, 1)$, $G_{n, p}$ is almost surely a Ramsey graph, as
\[
  \prob\bigl[\hom(G_{n, p}) \geq C\log n\bigr] \leq \binom{n}{C\log n}\left(p^{\binom{C\log n}{2}} + (1 - p)^{\binom{C\log n}{2}}\right) = o(1),
\]
for some $C = C(p) > 0$. 

Despite decades of intensive research, all known constructions of Ramsey graphs still rely on randomness, thus raising the belief that Ramsey graphs inherently ``random-like'' and share certain ``richness'' properties of dense random graphs. The first result that cemented this connection is due to Erd\H{o}s and Szemer\'{e}di, showing that Ramsey graphs are also dense.

\begin{theorem}[Erd\H{o}s--Szemer\'{e}di~\cite{erdos1972ramsey}]
  \label{thm:ramsey-density}

  Let $G$ be a graph on $n$ vertices with edge density $d \leq 1/2$. Then $G$ contains a homogeneous set of size greater than $K = \frac{c}{d\log(1/d)} \log n$, for some constant $c > 0$ and sufficiently large $n$.
\end{theorem}

\begin{proof}
  Suppose not. Pick $c < 1/32$. Since $G$ contains $d\binom{n}{2}$ edges, there exist $m \geq n/2$ vertices in $G$ with degree at most $dn$, by the pigeonhole principle. We work with the graph $G'$ induced by these $m$ vertices. Let $I$ be the largest independent set in $G'$ and let $k = |I|$. Since $\deg(v) \leq dn$ for all $v \in I$, we have $\sum_{v \in I} \deg(v) \leq dnk \leq 2dmk$. Put $s = 8dk$. Again by pigeonhole, there are at least $m/2 - k$ vertices $v \in V(G') \setminus I$ with $\deg_{I}(v) < s$. Let $J$ be the set of these vertices. Consider the family $\mathcal{F}$ of subsets of $I$ with size less than $s$. Partition $J$ by each vertex's neighbourhood in $I$, and note that each resulting class corresponds to an element of $\mathcal{F}$. By the standard bounds $\binom{x}{r} \leq \left(ex/r\right)^{r}$ and $1 + x \leq e^{x}$,
  \[
    |\mathcal{F}| \leq s\binom{k}{s - 1} \leq s\left(\frac{ek}{s - 1}\right)^{s - 1} \!\!= s\left(\frac{s}{s - 1}\right)^{s - 1}\left(\frac{ek}{s}\right)^{s - 1} \!\!\leq se\left(\frac{ek}{s}\right)^{s - 1} \!\!= \frac{s^2}{k}\left(\frac{ek}{s}\right)^s.
  \]
  By assumption $|I| = k < K$, so
  \[
    s < 8dK = \frac{8c\log n}{\log(1/d)} = O(\log n).
  \]
  Since $s/k = 8d$, 
  \[
    |\mathcal{F}| \leq \frac{s^2}{k}\left(\frac{ek}{s}\right)^s \leq O(\log n) \cdot \left(\frac{e}{8} \cdot \frac{1}{d}\right)^{s} \leq O(\log n) \cdot \left(\frac{1}{d}\right)^{\frac{8c\log n}{\log(1/d)}} = O(\log n) \cdot n^{8c} < n^{1/3}.
  \]
  But then 
  \[
    \frac{|J|}{|\mathcal{F}|} \geq \frac{m/2 - k}{n^{1/3}} \geq \sqrt{n},
  \]
  so by pigeonhole, there exists a class of vertices $Z$ in $J$ of size at least $\sqrt{n}$. Let $G'' = G[Z]$. Note that $G''$ does not contain an independent set of size $s$, as otherwise exchanging it with the corresponding subset of $I$ would create a larger independent set in $G'$, contradicting the maximality of $I$. Let $t$ be the size of the largest clique in $G''$, and note that $t < K$ by assumption. Erd\H{o}s--Szekeres theorem and the standard binomial bounds give
  \[
    \sqrt{n} \leq |Z| \leq \binom{s + K - 2}{s - 1} \leq \binom{s + K}{s} \leq \binom{8dK + K}{8dK} \leq \left(\frac{e(8d + 1)}{8d}\right)^{s} \leq \left(\frac{2}{d}\right)^s,
  \]
  where the last inequality holds because $e(1 + 8d)/8 \leq 5e/8 \leq 2$ for $d \leq 1/2$. But then since $d \leq 1/2$ and $c < 1/32$, 
  \[
    \left(\frac{2}{d}\right)^s \leq \left(\frac{2}{d}\right)^{\frac{8c\log n}{\log(1/d)}} = n^{8c \cdot \frac{\log(2/d)}{\log(1/d)}} = n^{8c\left( 1 + \frac{\log 2}{\log(1/d)}\right)} \leq n^{16c} < \sqrt{n},
  \]
  contradiction.
\end{proof}

Recall that the Erd\H{o}s--Szekeres theorem (Theorem \ref{thm:erdos-szekeres}) gives $\hom(G) \geq \frac{1}{2} \log_2 n$, regardless of the edge density. Theorem~\ref{thm:ramsey-density} improves this when the density is small. For example, if $d = 1/n^{\eps}$ for some $\eps > 0$, then $\hom(G) = \Omega\left(\frac{n^{\eps}}{\log n^{\eps}} \log n\right)$, vastly larger than $\frac{1}{2}\log_2 n$. Intuitively, a sparse graph should contain a large independent set, and the theorem makes this notion precise. 

If $G$ is $C$-Ramsey, then applying Theorem~\ref{thm:ramsey-density} to whichever $G$ or $\overline{G}$ has density $d \leq 1/2$ yields $c/(d\log(1/d)) \leq C$, forcing $d$ to be bounded away from $0$. The same argument applied to $\overline{G}$ forces $d$ to be bounded away from $1$. This gives the following corollary.

\begin{corollary}
  \label{cor:ramsey-density}
  For any $C$-Ramsey graph $G$, there exists $\delta = \delta(C) \in (0, 1/2)$ such that the edge density of $G$ lies in the interval $[\delta, 1-\delta]$.
\end{corollary}

Due to the rich variety of induced subgraphs in dense random graphs, a popular angle to study the pseudorandom properties of Ramsey graphs is to consider their induced subgraphs. Some notable results are a theorem of Shelah~\cite{shelah1997erdhosrenyiconjecture} which asserts that every $n$-vertex Ramsey graph contains $2^{\Omega(n)}$ non-isomorphic induced subgraphs, and a theorem of Pr\"{o}mel and R\"{o}dl~\cite{PROMEL1999379} which shows the existence of $\delta = \delta(C) > 0$ such that every $C$-Ramsey graph on $n$ vertices contains an induced copy of all graphs on at most $\delta \log n$ vertices. 

In this dissertation, we will focus on the edge statistics of induced subgraphs in Ramsey graphs. For a graph $G$, let 
\[
  \Phi(G) \coloneqq \bigl\{e(H) : H \text{ is an induced subgraph of } G\bigr\}.
\]
Erd\H{o}s and McKay~\cite{EFS} conjectured that for any Ramsey graph $G$ on $n$ vertices, there exists $\delta = \delta(C) > 0$ such that
\[
  \bigl\{1, 2, \ldots, \delta n^2\bigr\} \subseteq \Phi(G).
\]
Calkin, Frieze, and McKay~\cite{erdosmckayrandomgraph} proved it for random graphs, but the conjecture for Ramsey graphs remained open for decades. Significant progress began with Narayanan, Sahasrabudhe, and Tomon~\cite{narayanan2017ramseygraphsinducesubgraphs}, who established $|\Phi(G)| = n^{2 - o(1)}$ for Ramsey graph $G$ on $n$ vertices. Their proof builds on ideas from Bukh and Sudakov~\cite{BUKH2007612}, who showed that every Ramsey graph contains a linear sized subgraph with many distinct degrees, a feature shared with dense random graphs. Kwan and Sudakov~\cite{kwan2021ramseygraphsinducesubgraphs} later gave an optimal bound $|\Phi(G)| = \Omega(n^2)$ via a refined argument. Finally, Kwan, Sah, Sauermann, and Sawhney~\cite{Kwan_2023} closed the conjecture via Fourier-analytic arguments on graph polynomials.

This dissertation starts by studying the result of Bukh and Sudakov on large induced subgraphs in Ramsey graphs with many distinct degrees. This work not only tightens the connection between Ramsey graphs and random graphs, but also builds the foundation for the subsequent $|\Phi(G)| = \Omega(n^2)$ result by Kwan and Sudakov which we will present in the following chapter. Lastly, we outline the proof of the Erd\H{o}s--McKay conjecture by Kwan, Sah, Sauermann, and Sawhney.

\subsection{Anticoncentration}\label{sec:anticoncentration}

The core probabilistic mechanism in this dissertation is \textit{anticoncentration}. In the following chapters, we will repeatedly sample a uniformly random subset $U\subseteq V(G)$ and study the graph statistics of $G[U]$, such as vertex degrees or edge counts. The guiding principle is that if a random variable depends on many independent choices, then it should not concentrate too heavily on any  particular value.

For example, a question to ask in this setting is \textit{how likely it is for two vertices $x, y \in V(G)$ to have the same degree in $G[U]$?}

Observe that $\deg_U(x)-\deg_U(y)$ counts only vertices in $D = N_G(x)\triangle N_G(y)$. Write
$A = N(x)\setminus N(y), B = N(y)\setminus N(x)$. Then
\begin{equation}\label{eq:deg-diff}
  \deg_U(x) - \deg_U(y) = |U \cap A| - |U \cap B|,
\end{equation}
so $x, y$ have the same degree in $G[U]$ precisely when $|U \cap A|=|U \cap B|$. Since each $v \in A$ contributes $+1$ to the difference $|U \cap A| - |U \cap B|$ with probability $1/2$ and each $v \in B$ contributes $-1$ with probability $1/2$, the difference is a sum of $|D| = |A| + |B|$ independent random variables, which has standard deviation $\Theta(\sqrt{n})$ whenever $|D| \geq \eps n$ is large. Erd\H{o}s--Littlewood--Offord then gives $\prob[\deg_U(x) = \deg_U(y)] = O(1/\sqrt{n})$, and a \textit{collision graph} construction with Tur\'{a}n's theorem yields many distinct degrees in $U$.

An intuition for this $O(1/\sqrt{n})$ bound is that $\deg_U(x) - \deg_U(y)$ is a \textit{linear} polynomial of many independent indicators, which is approximately normally distributed with standard deviation $\Theta(\sqrt{n})$. Since the probability density function $\phi(x)$ of a normal distribution with standard deviation $\sigma$ is at most $1/(\sigma\sqrt{2\pi})$ for all $x$, the probability that $\deg_U(x) - \deg_U(y)$ equals any particular value is $O(1/\sigma) = O(1/\sqrt{n})$, and this anticoncentration bound is the key to the results in Chapters~\ref{sec:ch2} and~\ref{sec:ch3}. 

In Chapter~\ref{sec:ch4}, we will study the induced edge count $e(U)$ itself, which is instead a \textit{quadratic} polynomial of indicators. It turns out a central limit theorem still holds for $e(U)$ at the large scale with standard deviation $\sigma(e(U)) = \Theta(n^{3/2})$, and we prove its point probabilities follow the $O(1/\sigma) = O(n^{-3/2})$ bound that a Gaussian would give, even though $e(U)$ need not be Gaussian at the small scale. This is the same anticoncentration principle as in Chapters~\ref{sec:ch2} and \ref{sec:ch3}, sharpened to a statement about $e(U)$ rather than a degree difference.

\subsection{Notations and tools}

For $n \in \N$, write $[n] \coloneqq \{1, 2, \ldots, n\}$. For set $X$, write $\binom{X}{2}$ as the set of unordered pairs of elements in $X$. Denote the union of sets $X, Y$ as $X \sqcup Y$ if $X \cap Y = \emptyset$.

Given graph $G$ on $n$ vertices, we call $e(G) \coloneqq |E(G)|$ the \textit{size} of $G$, and the ratio 
\[
  d(G) \coloneqq e(G)\binom{n}{2}^{-1}
\]
the \textit{density} of $G$. For vertex $v \in V(G)$, denote $N(v) \coloneqq \{w \in V(G) : \{v, w\} \in E(G)\}$ as the \textit{neighbourhood} of $v$ and $\deg(v) \coloneqq |N(v)|$ as the \textit{degree} of $v$. Write $\overline{G}$ as the \textit{complement graph} of $G$ and $\overline{N(v)} \coloneqq V(G) \setminus N(v)$ as the \textit{non-neighbourhood} of $v$. For $S \subseteq V(G)$, denote $G - S$ as the subgraph of $G$ obtained by deleting $S$ and all edges incident to $S$. For $U \subseteq V(G)$, we call $G[U] \coloneqq G - (V(G) \setminus U)$ the $\textit{induced subgraph}$ of $U$ in $G$. Write $e(U) \coloneqq e(G[U])$ and $d(U) \coloneqq d(G[U])$ as the size and density of $U$, respectively. Write $N_U(v) \coloneqq N(v) \cap U$, $\deg_U(v) \coloneqq |N_U(v)|$, and $\overline{N_U(v)} \coloneqq U \setminus N_U(v)$ as the neighbourhood, degree, and non-neighbourhood of $v$ in $U$, respectively. For $A, B \subseteq V(G)$, write $e(A, B) \coloneqq e(A \cup B) - e(A) - e(B)$ and
\[
  d(A, B) \coloneqq \frac{e(A, B)}{|A||B|}
\]
as the edge count and density between $A$ and $B$, respectively.

For functions $f, g: \R \to \R$, we use the standard asymptotic notations:
\begin{itemize}
\item $f = O(g)$ and $f \ll g$ mean there exists constant $C > 0$ such that $f \leq C \cdot g$.
\item $f = \Omega(g)$ and $f \gg g$ mean there exists constant $c > 0$ such that $f \geq c \cdot g$.
\item $f = \Theta(g)$ means $f = O(g)$ and $f = \Omega(g)$.
\item $f = o(g)$ means $f/g \to 0$.
\item $f = \omega(g)$ means $f/g \to \infty$.
\end{itemize}

The results in this dissertation mostly concern large graphs, so we omit floor and ceiling signs when they are not crucial.

We call subset $U \subseteq X$ \textit{uniformly random} if each $x \in X$ is in $U$ independently with probability $1/2$. For event $A$, write its indicator function as
\[
  \mathbbm{1}_A \coloneqq \begin{cases}
    1 & \text{if } A \text{ occurs}, \\
    0 & \text{otherwise}.
  \end{cases}
\]


We will use the following probability bounds throughout the dissertation.

\begin{proposition}[Markov's inequality]
  \label{prop:markov}
  Let $X$ be a non-negative random variable. Then for $t \geq 0$,
  \[
    \prob\bigl[X \geq t\bigr] \leq \frac{\E[X]}{t}.
  \]
  \qed
\end{proposition}

% \begin{proof}
%   Since $t \cdot \mathbbm{1}_{X \geq t} \leq X$, taking expectations gives the result.
% \end{proof}

\begin{proposition}[Chebyshev's inequality]
  \label{prop:chebyshev}

  Let $X$ be a random variable. Then for $t > 0$,
  \[
    \prob\bigl[|X - \E[X]| \geq t\bigr] \leq \frac{\Var(X)}{t^2}.
  \]
  \qed
\end{proposition}

% \begin{proof}
%   Applying Markov's inequality to $Y = (X - \E[X])^2$ gives
%   \[
%     \prob\bigl[|X - \E[X]| \geq t\bigr] = \prob\bigl[Y \geq t^2\bigr] \leq \frac{\E[Y]}{t^2} = \frac{\Var(X)}{t^2}.
%   \]
% \end{proof}

For $X = \mathbbm{1}_{A_1} + \cdots + \mathbbm{1}_{A_k}$, we often use the following inequality to bound the variance:
\begin{align*}
  \Var(X) 
  &= \E[X^2] - \E[X]^2 \\
  &= \sum_{i}\sum_j \prob[A_i \cap A_j] - \prob[A_i] \prob[A_j] \\
  &= \sum_i \bigl[\prob[A_i] - \prob[A_i]^2\bigr] + \sum_i \sum_{j \sim i} \bigl[\prob[A_i \cap A_j] - \prob[A_i] \prob[A_j]\bigr] \\
  &\leq \E[X] + \sum_i \sum_{j \sim i} \prob[A_i \cap A_j]
\end{align*}
where we write $i \sim j$ if $i \neq j$ and $A_i, A_j$ are dependent.

\begin{proposition}[Chernoff bound]
  \label{prop:chernoff}
  Let $X$ be a binomial random variable with parameters $n$ and $p$. Then for $t \geq 0$,
  \[
    \prob\bigl[X \leq \E[X] - t\bigr], \, \prob\bigl[X \geq \E[X] + t\bigr] \leq \exp\bigl(-2t^2/n\bigr).
  \]
  \qed
\end{proposition}

\begin{proposition}[Hoeffding's inequality]
  \label{prop:azuma-hoeffding}
  
  Let $X_1, X_2, \ldots, X_n$ be independent random variables with $0 \leq X_i \leq 1$ for each $i$, and write $X = \sum_{i = 1}^n X_i$. Then for $t \geq 0$, 
  \[
    \prob\bigl[|X - \E[X]| \geq t\bigr] \leq 2e^{-2t^2/n}.
  \]
  \qed
\end{proposition}

Many random variables turn out to be linear polynomials of independent Bernoulli random variables, to which Erd\H{o}s--Littlewood--Offord can be applied to show anticoncentration. We refer to these random variables as \textit{Littlewood--Offord type}.

\begin{definition}
  \label{def:littlewood-offord-type}
  A random variable $X$ is of \textit{$(n, p)$-Littlewood--Offord type} if it is of the form $X = a_1 \xi_1 + a_2 \xi_2 + \cdots + a_n \xi_n + C$, where $a_1, a_2, \ldots, a_n \in \Z \setminus \{0\}$ and $C \in \Z$ are fixed and $\xi_1, \xi_2, \ldots, \xi_n$ are independent Bernoulli random variables with parameter $p$.
\end{definition}

\begin{proposition}[Erd\H{o}s--Littlewood--Offord theorem]
  \label{prop:littlewood-offord}
  Let $p > 0$ be constant and $X$ be a random variable of $(n, p)$-Littlewood--Offord type. Then for any $t \in \Z$, 
  \[
    \prob[X = t] = O\bigl(1/\sqrt{n}\bigr).
  \]
\end{proposition}

\begin{proof}
  First consider the case where all coefficients $a_i$ are positive. Without loss of generality, assume $X \coloneqq \sum_{i = 1}^n a_i \xi_i$. For $t \in \Z$, let
  \[
    \mathcal{A}_t \coloneqq \left\{A : \sum_{i \in A} a_i = t\right\} \subseteq [n].
  \]
  Since $a_i \geq 1$ for all $i$, we have $1 + \sum_{i \in A} a_i \leq \sum_{i \in B} a_i$ whenever $A \subset B$. Thus, no elements of $\mathcal{A}_t$ are contained in another. Consider the chain $C = (S_1, S_2, \ldots, S_n)$ of $n$ subsets of $[n]$ such that $\emptyset = S_1 \subset S_2 \subset \cdots \subset S_n = [n]$. There are $n!$ such chains $C$ in $[n]$, and for $A \subseteq [n]$, there are $|A|!(n - |A|)!$ of them containing $A$. Since no chain can contain multiple elements of $\mathcal{A}_t$,
  \[
    \sum_{A \in \mathcal{A}_t} |A|!(n - |A|)! \leq n! \implies \sum_{A \in \mathcal{A}_t} \frac{1}{\binom{n}{|A|}} \leq 1.
  \] 
  By Stirling's formula on binomial coefficients,
  \begin{align*}
    \prob\bigl[X = t\bigr] 
    &= \sum_{A \in \mathcal{A}_t} p^{|A|}(1 - p)^{n - |A|} \\
    &\leq \sum_{A \in \mathcal{A}_t} \frac{1}{\binom{n}{|A|}} \cdot \binom{n}{|A|}p^{|A|}(1 - p)^{n - |A|} \\
    &\leq \left(\sum_{A \in \mathcal{A}_t} \frac{1}{\binom{n}{|A|}}\right) \cdot \binom{n}{np}p^{np}(1 - p)^{n - np} \\
    &\leq \frac{1 + o(1)}{\sqrt{2\pi p(1 - p)n}} = O\bigl(1/\sqrt{n}\bigr).
  \end{align*}

  Now consider the general case. Let $I \coloneqq \{i : a_i \geq 1\}$ and $J \coloneqq \{i : a_i \leq -1\}$. We may assume $|I| \geq n/2$ by passing to $-X$ and swapping $I$ with $J$ if necessary. Conditioning on $(\xi_j)_{j \in J}$, $X$ is of $(|I|, p)$-Littlewood--Offord type with positive coefficients, so
  \[
    \prob\bigl[X = t \mid (\xi_j)_{j \in J}\bigr] = O\bigl(1/\sqrt{|I|}\bigr) = O\bigl(1/\sqrt{n}\bigr).
  \]
  The law of total probability gives
  \[
    \prob\bigl[X = t\bigr] = \E\left[\prob\bigl[X = t \mid (\xi_j)_{j \in J}\bigr]\right] = O\bigl(1/\sqrt{n}\bigr).
  \]
\end{proof}

We also use the following version of Tur\'{a}n's theorem.

\begin{proposition}[Tur\'{a}n's theorem]
  \label{prop:turan}
  
  Every graph $G$ with $n$ vertices and $m$ edges contains an independent set of size at least $n^2/(2m + n)$.
\end{proposition}

\begin{proof}
  Consider a random ordering $<$ on $V(G)$. Define 
  \[
    I \coloneqq \bigl\{v \in V : v < u, \, \forall u \in N(v)\bigr\}.
  \]
  For distinct $u, v \in V(G)$, either $u > v$ or $v > u$, so $u, v$ cannot both be in $I$ if they are adjacent. Thus $I$ is an independent set. Since $v \in I$ if and only if $v$ is the least element among $v$ and its neighbours, 
  \[
    \E\bigl[|I|\bigr] = \sum_{v \in V} \prob\bigl[v \in I\bigr] = \sum_{v \in V} \frac{1}{1 + \deg(v)}.
  \]
  The function $f(x) = 1/(x + 1)$ is convex on $[0, \infty)$, so by Jensen's inequality, $G$ contains an independent set $I$ of size 
  \[
    |I| \geq \sum_{v \in V} \frac{1}{1 + \deg(v)} \geq \frac{n}{1 + \frac{1}{n}\sum_{v \in V} \deg(v)} = \frac{n^2}{2m + n}.
  \]
\end{proof}