\section{Subgraphs with Many Degrees}
\label{sec:ch2}

In this chapter, we present the following result of Bukh and Sudakov.

\begin{theorem}[Bukh--Sudakov {\cite{BUKH2007612}}]
  \label{thm:bs2007}
  Every Ramsey graph on $n$ vertices induces a subgraph on $\Omega(n)$ vertices with $\Omega(\sqrt{n})$ distinct degrees.
\end{theorem}

This was originally conjectured by Erd\H{o}s, Faudree, and S\'{o}s \cite{EFS} in light of the observation that the random graph $G_{n, 1/2}$ exhibits such a property almost surely. In particular, this theorem already has implications for the edge statistics of Ramsey graphs $G$, as the subgraphs $G - \{v\}$ for $v \in V(G)$ have $\Omega(\sqrt{n})$ different sizes. This idea lays the foundation for the proof in the next chapter.

To prove Theorem \ref{thm:bs2007}, Bukh and Sudakov introduced the notion of \textit{diversity}, which captures the idea that the neighbourhoods of most pairs of vertices are not too similar.

% , but the $\Omega(\sqrt{n})$ bound is in fact not optimal for both Ramsey graphs and $G_{n, 1/2}$. 

% In particular, an upper bound of $O(n^{2/3})$ is noted in \cite{BUKH2007612}, which was later proved to be the correct order for Ramsey graphs by Jenssen, Keevash, Long, and Yepremyan \cite{jenssen2019distinctdegreesinducedsubgraphs} via a continuous relaxation approach. The corresponding order for $G_{n, 1/2}$ was also matched in an unpublished work of Conlon, Morris, Samotij, and Saxton \cite{conlon2019distinctdegrees}, further reinforcing the case that Ramsey graphs resemble random graphs in various respects. We will, however, focus on the original result of Bukh and Sudakov, as it introduces a key characteristic of Ramsey graphs that will be insightful in the following chapter.

\begin{definition}
  \label{def:diversity}

  An $n$-vertex graph $G$ is \textit{$(\eps, \delta)$-diverse} if for $x \in V(G)$, we have $|N(x) \triangle N(y)| \geq \eps n$ for $\leq n^{\delta}$ vertices $y \in V(G)$.
\end{definition}

Notice that $(\eps, \delta)$-diversity directly conditions our assumption that $|N(x) \triangle N(y)| = D \geq \eps n$ in the discussion of the anticoncentration principle in Section \ref{sec:anticoncentration}. In particular, we may show in a diverse graph that $\prob[\deg_U(x) = \deg_U(y)] = O(1/\sqrt{n})$ for all but $\leq n^{\delta}$ pairs of vertices $x, y \in V(G)$, from which we use Jensen's inequality to conclude that there are many distinct degrees in a diverse graph. The following lemma formalizes this observation and reduces Theorem \ref{thm:bs2007} to showing every Ramsey graph contains a diverse subgraph of linear size.

\begin{lemma}
  \label{lem:diversity-degrees}

  Let $\delta \in (0, 1/2)$ and let $G$ be an $(\eps, \delta)$-diverse graph on $n$ vertices for large enough $n$. Then for $n/4 \leq m \leq 3n/4$, there is a subgraph on $m$ vertices containing vertices of $\sqrt{\eps n}/10$ distinct degrees.
\end{lemma}

\begin{proof}
  Fix $m$ as above. Let $U$ be chosen uniformly at random among all subsets of $V(G)$ of size $m$. It suffices to show that with positive probability, $G[U]$ contains $\Omega(\sqrt{n})$ vertices of distinct degrees.

  For distinct $x, y \in V(G)$, define $\mathbbm{1}_{xy}$ to be the indicator for the event that $x, y \in V(H)$ and $\deg_U(x) = \deg_U(y)$. Then $\E[\mathbbm{1}_{xy}] \leq \prob[\deg_U(x) = \deg_U(y)]$. Put $D = N_G(x) \triangle N_G(y)$, and $k = |D \cap U|$. Write $A = N_G(x) \setminus N_G(y)$ and $B = N_G(y) \setminus N_G(x)$. Then 
  \[
    \prob[\deg_U(x) = \deg_U(y)] = \prob\left[|A \cap U| = |B \cap U|\right].
  \]
  Suppose $|D| \geq \eps n$. Since $\frac{|D|}{4} \leq \frac{m|D|}{n} \leq \frac{3|D|}{4}$, Proposition \ref{prop:hypergeometric} yields
  \[
    \prob\left[k < \frac{|D|}{8} \text{ or } k > \frac{7|D|}{8}\right] \leq \prob\left[\left|k - \E[k]\right| > \frac{|D|}{8}\right] \leq e^{-\Theta(|D|)} = e^{-\Theta(n)} = o(n^{-1/2}).  
  \]
  On the other hand, recall that $U$ is a uniformly random subset of a fixed size $m$. Since every $k$-subset of $D$ is equally likely to be $U \cap D$ by symmetry, conditioning on $|U \cap D| = k$, the $k$ elements of $U$ that land in $D$ form a uniformly random subset of $D$ of size $k$. Conditioning on $k$, the event $|A \cap U| = |B \cap U| = k/2$ then follows the hypergeometric distribution $\mathrm{Hyp}(|D|, |A|, k)$. Thus,
  \[
    \prob\left[\deg_U(x) = \deg_U(y) \, \middle| \, \frac{|D|}{8} \leq k \leq \frac{7|D|}{8}\right] = \frac{\binom{|A|}{k/2}\binom{|B|}{k/2}}{\binom{|D|}{k}}.
  \]
  Since $|A| + |B| = |D|$, by the log-concavity of the binomial coefficients we have $\binom{|A|}{k/2}\binom{|B|}{k/2} \leq \binom{|D|/2}{k/2}^2$. It now follows from Stirling's formula and $k(|D| - k) \geq \frac{7}{64}|D|^2$ that the above probability does not exceed
  \begin{equation}
    \frac{\binom{|D|/2}{k/2}^2}{\binom{|D|}{k}} = (1 + o(1))\sqrt{\frac{2|D|}{\pi k(|D| - k)}} \leq \frac{3}{\sqrt{|D|}} \leq \frac{3}{\sqrt{\eps n}}. \label{eq:prob-deg-diff}
  \end{equation}
  Thus if $|N_G(x) \triangle N_G(y)| \geq \eps n$ then $\E[\mathbbm{1}_{xy}] \leq 3/\sqrt{\eps n} + o(n^{-1/2}) \leq 4/\sqrt{\eps n}$.

  Now consider the degree collision graph $H$ on $U$, where $x, y \in U$ are adjacent if and only if $\deg_U(x) = \deg_U(y)$. Since $G$ is $(\eps, \delta)$-diverse, $\E[\mathbbm{1}_{xy}] \leq 4/\sqrt{\eps n}$ for all but $\leq n^{1 + \delta} < n^{3/2}$ pairs $x, y \in V(G)$. Hence, the expected number of edges in $H$ is
  \[
    \E[e(H)] = \sum_{x, y \in V(G)} \E[\mathbbm{1}_{xy}] \leq n^{1 + \delta} + \frac{4}{\sqrt{\eps n}} \cdot n^2 \leq \frac{5}{\sqrt{\eps}}n^{3/2}.
  \] 
  It now follows from Turan's Theorem (Proposition \ref{prop:turan}) that $H$ contains an independent set $I \subseteq U$ of size at least 
  \[
    |I| \geq \frac{n^2}{2\E[e(H)] + n} \leq \frac{\sqrt{\eps n}}{10}.
  \]
  In particular, each vertex in $I$ has a distinct degree in $G[U]$.
\end{proof}

% \textit{Remark.} The hypergeometric distribution $\mathrm{Hyp}(|D|, |A|, k)$ has variance 
% \[
%   \sigma^2 = \frac{k|A||B|(|D| - k)}{|D|^2(D - 1)} \leq \frac{k(|D| - k)}{4(|D| - 1)},
% \]
% where inequality follows from AM-GM. When $k$ is not the tail end of the distribution, we have $\sigma = \Theta(\sqrt{D}) = \Theta(\sqrt{n})$, and the Stirling bound in \eqref{eq:prob-deg-diff} is asymptotically $1/(\sigma\sqrt{2\pi})$ as $n \to \infty$, which is exactly the maximum density of a normal distribution with standard deviation $\sigma$. This confirms the intuition from Section~\ref{sec:anticoncentration} that the degree difference is approximately normally distributed so its point probabilities are $O(1/\sigma) = O(1/\sqrt{n})$.

It remains to show that every Ramsey graph contains a diverse subgraph of linear size. Instead of directly proving this, we prove the same statement on a strengthening of diversity called \textit{richness}.

\begin{definition}
  \label{def:richness}
  An $n$-vertex graph $G$ is \textit{$(\eps, \delta)$-rich} if, for $W \subseteq V(G)$ with $|W| \geq \delta n$, we have $|N(v) \cap W| \geq \eps|W|$ and $|\overline{N(v)} \cap W| \geq \eps|W|$ for $\leq n^{\delta}$ vertices $v \in V(G)$.
\end{definition}

The key property of richness is that it ensures the intersection of vertex neighbourhoods with arbitrary large vertex sets are not extremely large or small. When $W$ is the neighborhood of another vertex, richness ensures both sides of the symmetric difference $|N(x) \triangle N(y)|$ are large, rather than merely the symmetric difference itself being large. This is why richness is a stronger condition than diversity.

\begin{lemma}
  \label{lem:richness-implies-diversity}
  Let an $n$-vertex graph $G$ be $(\eps, \delta)$-rich, with $\delta \leq 1/2$. Then $G$ is $(\eps/2, \delta)$-diverse.
\end{lemma}

\begin{proof}
  Note that for each $x \in V(G)$, either $|N(x)| \geq n/2$ or $|\overline{N(x)}| \geq n/2$. If $|N(x)| \geq n/2$, then richness implies that
  \[
    |\overline{N(y)} \cap N(x)| \geq \eps |N(x)| \geq \eps n/2
  \]
  for $\leq n^{\delta}$ vertices $y \in V(G)$. Similarly, if $|\overline{N(x)}| \geq n/2$, then
  \[
    |N(y) \cap \overline{N(x)}| \geq \eps |\overline{N(x)}| \geq \eps n/2
  \]
  for $\leq n^{\delta}$ vertices $y \in V(G)$. But then in either case,
  \[
    |N(x) \triangle N(y)| = |\overline{N(y)} \cap N(x)|  + |N(y) \cap \overline{N(x)}| \geq \eps n/2.
  \]
\end{proof}

We now show that every Ramsey graph contains a rich subgraph of linear size.

\begin{lemma}
  \label{lem:rich-subgraph}

  For $\delta \in (0, 1)$, there exists $\eps = \eps(C, \delta) > 0$ such that every $C$-Ramsey graph $G$ on $n$ vertices contains an $(\eps, \delta)$-rich subgraph on $\Omega(n)$ vertices.
\end{lemma}

\begin{proof}
  Suppose not. Let $K = K(C)$ be a large constant to be determined. Let $c = 4^{-K}$ and $\eps = 1/K$. We inductively construct a sequence of disjoint vertex sets $S_1, S_2, \ldots, S_K$ of size $s = (cn)^\delta/2$, such that for all $S_i$ one of the following conditions holds:
  \begin{enumerate}
    \item[(i)] $d(S_i, S_j) < 4\eps$ for $j > i$,
    \item[(ii)] $d(S_i, S_j) > 1 - 4\eps$ for $j > i$.
  \end{enumerate}
  Alongside, we also construct a sequence of induced subgraphs $G = G[U_0] \supset G[U_1] \supset \cdots \supset G[U_K]$, such that $|U_i| \geq (\delta/4)|U_{i - 1}|$ and $S_i \subseteq U_{i - 1}$ for all $i$.

  We first show how this construction gives the lemma. By working with $\overline{G}$ instead of $G$ if necessary, we may assume that $r \geq K/2$ of the sets $S_{1}, \ldots, S_{K}$ satisfy condition (i), say $S_{i_1}, \ldots, S_{i_r}$. Put $S = \bigsqcup_{k = 1}^r S_{i_k}$. Since each $S_{i_k}$ has size $s$ and $\eps = 1/K$,
  \[
    d(S) = \frac{\sum_{k = 1}^r e(S_{i_k})}{\binom{|S|}{2}} + \frac{\sum_{k < l} e(S_{i_k}, S_{i_l})}{\binom{|S|}{2}} \leq \frac{r\binom{s}{2}}{\binom{rs}{2}} + \max_{(k, l)} d(S_{i_k}, S_{i_l}) \leq \frac{1}{r} + 4\eps \leq \frac{6}{K}.
  \]
  The Erd\H{o}s--Szemer\'{e}di theorem (Theorem \ref{thm:ramsey-density}) now yields a homogeneous set $A \subseteq S$ of size
  \[
    |A| \geq \frac{\gamma K}{10\log(K/10)} \log |S| = \frac{\gamma \delta K}{10\log(K/10)} \log cn = \frac{\gamma \delta K}{10\log(K/10)} \log n + o(\log n),
  \]
  for some absolute constant $\gamma > 0$ given by the theorem. Take $K$ to be large enough so that $\frac{\gamma \delta K}{10\log(K/10)} > C$. Then $|A| > C\log n$, contradiction.

  We now construct the sets $S_i$ and $U_i$ inductively. Suppose $U_1, \ldots, U_{i - 1}$ and $S_1, \ldots, S_{i - 1}$ have been constructed. By induction, $|U_{i - 1}| \geq (\delta/4)^{i - 1}|U_0| \geq cn$. Since $G[U_{i - 1}]$ is not $(\eps, \delta)$-rich by assumption, $U_{i - 1}$ contains subsets $W_i, Y_i$ with $|W_i| \geq \delta|U_{i - 1}|$ and $|Y_i| = (cn)^{\delta}$, such that for each $y \in Y_i$ we have $|N_{G_{i - 1}}(y) \cap W_i| < \eps|W_i|$ or $|\overline{N_{G_{i - 1}}(y)} \cap W_i| < \eps|W_i|$. We consider two cases.

  Suppose $|N_{G_{i - 1}}(y) \cap W_i| < \eps|W_i|$ for half the vertices $y \in Y_i$, and let $S_i$ be the corresponding subset of $Y_i$, so $|S_i| = (cn)^{\delta}/2$. Put $T_i = W_i \setminus S_i$ and note that $|T_i| \geq |W_i|/2$, as $|S_i| \leq |W_i|/2$ for large enough $n$. Then for $y \in S_i$,
  \[
    d(y, T_i) = \frac{|N_{G_{i - 1}}(y) \cap T_i|}{|T_i|} \leq \frac{|N_{G_{i - 1}}(y) \cap W_i|}{|T_i|} < \frac{\eps |W_i|}{|W_i|/2} = 2\eps.
  \]
  Partition $T_i$ into $U_i \sqcup Z_i$ where $U_i = \{v \in T_i : d(v, S_i) \leq 4 \eps\}$. We show that $|U_i| \geq (\delta/4)|U_{i - 1}|$ by double counting $e(T_i, S_i)/|S_i|$. Since $d(v, S_i) > 4\eps$ for $v \in Z_i$,
  \[
    \frac{e(T_i, S_i)}{|S_i|} = \sum_{v \in T_i} d(v, S_i) \geq \sum_{v \in Z_i} d(v, S_i) > 4\eps|Z_i|.
  \]
  On the other hand, since $d(y, T_i) < 2\eps$ for $y \in S_i$,
  \[
    \frac{e(T_i, S_i)}{|S_i|} = \frac{1}{|S_i|}\sum_{y \in S_i} e(y, T_i) = \frac{|T_i|}{|S_i|}\sum_{y \in S_i} d(y, T_i) < 2\eps|T_i|.
  \]
  Combining both inequalities yields $|Z_i| < |T_i|/2$. Therefore
  \[
    |U_i| = |T_i| - |Z_i| > \frac{|T_i|}{2} \geq \frac{|W_i|}{4} \geq \frac{\delta}{4}|U_{i - 1}|.
  \]
  For $j > i$, since $S_j \subseteq U_j \subseteq U_i$, we have $d(v, S_i) \leq 4\eps$ for every $v \in S_j$, and averaging gives $d(S_i, S_j) \leq 4\eps$. Thus $S_i$ satisfies condition~(i).

  Otherwise, $|\overline{N_{G_{i - 1}}(y)} \cap W_i| < \eps|W_i|$ for half the vertices $y \in Y_i$. Let $S_i$ be the corresponding subset of $Y_i$ and put $T_i = W_i \setminus S_i$, so again $|T_i| \geq |W_i|/2$ and $|S_i| = (cn)^{\delta}/2$. For $y \in S_i$,
  \[
    1 - d(y, T_i) = \frac{|\overline{N_{G_{i - 1}}(y)} \cap T_i|}{|T_i|} \leq \frac{|\overline{N_{G_{i - 1}}(y)} \cap W_i|}{|T_i|} < \frac{\eps |W_i|}{|W_i|/2} = 2\eps,
  \]
  so $d(y, T_i) > 1 - 2\eps$ for every $y \in S_i$. Partition $T_i$ into $U_i \sqcup Z_i$ where $U_i = \{v \in T_i : d(v, S_i) \geq 1 - 4\eps\}$. We double count $|T_i| - e(T_i, S_i)/|S_i| = \sum_{v \in T_i}(1 - d(v, S_i))$. Since $1 - d(v, S_i) > 4\eps$ for $v \in Z_i$,
  \[
    |T_i| - \frac{e(T_i, S_i)}{|S_i|} = \sum_{v \in T_i}(1 - d(v, S_i)) \geq \sum_{v \in Z_i}(1 - d(v, S_i)) > 4\eps|Z_i|.
  \]
  On the other hand, using $d(y, T_i) > 1 - 2\eps$ for $y \in S_i$,
  \[
    |T_i| - \frac{e(T_i, S_i)}{|S_i|} = |T_i| - \frac{|T_i|}{|S_i|}\sum_{y \in S_i} d(y, T_i) < |T_i| - (1 - 2\eps)|T_i| = 2\eps|T_i|.
  \]
  Combining both inequalities gives $|Z_i| < |T_i|/2$, hence $|U_i| > |T_i|/2 \geq (\delta/4)|U_{i - 1}|$ as in the other case. For  $j > i$, each $v \in S_j \subseteq U_i$ satisfies $d(v, S_i) \geq 1 - 4\eps$, so $d(S_i, S_j) \geq 1 - 4\eps$. Thus $S_i$ satisfies condition~(ii). This completes the induction.
\end{proof}

Theorem \ref{thm:bs2007} now follows by a simple application of Lemmas \ref{lem:rich-subgraph}, \ref{lem:richness-implies-diversity}, and \ref{lem:diversity-degrees} in sequence.

\paragraph{Proof of Theorem \ref{thm:bs2007}.} Let $G$ be a $C$-Ramsey graph on $n$ vertices, and fix any $\delta \in (0, 1/2)$. By Lemma \ref{lem:rich-subgraph}, there exists $\eps = \eps(C, \delta) > 0$ and an $(\eps, \delta)$-rich induced subgraph $G[V'] \subseteq G$ with $|V'| \geq c n$ for some $c = c(C, \delta) > 0$. Put $n' = |V'|$. By Lemma \ref{lem:richness-implies-diversity}, $G[V']$ is $(\eps/2, \delta)$-diverse on $n'$ vertices. Applying Lemma \ref{lem:diversity-degrees} to $G[V']$ with $m = n'/2$, we obtain an induced subgraph of $G[V']$ on $m \geq cn/2 = \Omega(n)$ vertices containing $\frac{\sqrt{\eps n'/2}}{10} \geq \frac{\sqrt{\eps c n/2}}{10} = \Omega(\sqrt{n})$ vertices of distinct degrees. \qed